---
title: "This is my portfolio"
author: "Qi Draaisma"
date: "10-2-2021"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

``` {r echo=FALSE, message=FALSE}
library(tidyverse)
library(spotifyr)
library(compmus)
library(plotly)
```

### The saddest Rock 'n Roll track in the corpus

#### Cepstogram

``` {r}
Crying <-
  get_tidy_audio_analysis("6eLL7QTdMWdhhG4i3jHDR9") %>% # Change URI.
  compmus_align(bars, segments) %>%                     # Change `bars`
  select(bars) %>%                                      #   in all three
  unnest(bars) %>%                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  )

Crying %>%
  compmus_gather_timbre() %>%
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = basis,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  scale_fill_viridis_c() +                              
  theme_classic()
```

#### Self Similarity Matrix

``` {r}
Crying %>%
  compmus_self_similarity(pitches, "cosine") %>% 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")
```

***
In the two plots below, the biggest outlier of the corpus is represented with a cepstogram and a self similarity matrix (SSM) regarding pitches. The song is "Crying" by Roy Orbison. What can be extracted from both plots is that the song has a strong ending of some sort. In the cepstogram one sees a short rise in magnitude of the c02 vector, which means a certain aspect of the timbre is extra apparent at that moment. This is also seen within the SSM as two vertical and two horizontal lines, this means that near the end there is a bigger difference regarding the other parts of the song. When listening to the song, it becomes clear what this change is. Throughout the song, Orbison sings a sad song. However, in the end the singer reaches out for a strong high note. This is what causes the sudden change within the two plots.

### Introduction to my research on the evolution of Rock music
Course portfolio for Computational Musicology

For this portfolio, I will be looking at the evolution of the Rock genre. Specifically, from Rock 'n Roll into Rock music and eventually into more alternative Rock subgenres,
 like Grunge, Metal and alt Rock. For the Rock 'n Roll songs, the Spotify playlist "50's and 60's Rock n Roll" playlist has been chosen to process. There are 74 songs within 
 the playlist and includes classic Rock 'n Roll artists, like Elvis Presley, Chuck Berry, Little Richard and much more. The playlist for the Rock music will be the Rock Classics 
 playlist from Spotify itself. This playlist includes 145 songs of classic Rock songs and is very likely to be suited for this era of Rock music, because it contains a lot of 
 different styles of Rock within that era. Finally, for the alternative Rock subgenres a playlist named "90s Grunge/Rock/Punk-pop" is used in combination with the "Old school
 Metal" playlist with a total of 293 songs in both playlists. The playlists are used together, because the "90s Grunge/Rock/Punk-pop" playlist contains a lot of subgenres but 
 doesn't contain a lot of Metal, which was definitely apparent in this era of Rock music. 

 The decision to look at the evolution of Rock music came to me, because of my obsession with why people seem to know these classical Rock songs no matter the age of 
 the person. Classic Rock music seems to be something almost everyone can enjoy, albeit not all the subgenres. Within the alternative Rock category, there are bands like 
 The Offspring and Green Day, which are well known for their heavier "Punk" style Rock. Take for example "Basket Case" from Green Day, which has more than half a billion
 streams. This is not even accounting for the amount of times the song is listened to in any other way. Next to this Punk style, the genre still contains bands like U2 as well. 
 The style of U2 is completely different, but rooted from the same type of music. Rock.

 All in all, I expect the loudness of Rock music in it's evolution to rise, this because of the fact that in later stages, the guitars became more distorted. Next to that, I expect 
 that the valence, which measures the positivity of a song, will be lower in the alternative Rock category because of the fact that the songs are written in less of a "swing"
 than the early Rock 'n Roll days.

### Analyzing the energy and mood of 50s, 60s Rock 'n Roll days. 

```{r echo=FALSE, message=FALSE}

spotifyr::get_spotify_access_token() 

RockNRoll <- get_playlist_audio_features("", "2ii0K7mNdB89XgcOCUM64t")
RockClassics <- get_playlist_audio_features("", "37i9dQZF1DWXRqgorJj26U")
RockSubGenres <- get_playlist_audio_features("", "5dZcUWhn9DP8oOWVQVYLq4")

Rock <- bind_rows(
  RockNRoll %>% mutate(era = "Rock 'n Roll (50s, 60s)"),
  RockClassics %>% mutate(era = "Rock Classics (70s, 80s)"),
  RockSubGenres %>% mutate(era = "Rock Subgenres (90s, 00s)")
)

RockMeans <- bind_rows(
  RockNRoll %>% summarise(
    mean_valence = mean(valence),
    mean_danceability = mean(danceability),
    mean_energy = mean(energy),
    era = "Rock 'n Roll (50s, 60s)"
  ),
  RockClassics %>% summarise(
    mean_valence = mean(valence),
    mean_danceability = mean(danceability),
    mean_energy = mean(energy),
    era = "Rock Classics (70s, 80s)"
  ),
  RockSubGenres %>% summarise(
    mean_valence = mean(valence),
    mean_danceability = mean(danceability),
    mean_energy = mean(energy),
    era = "Rock Subgenres (90s, 00s)"
  ),
  )

rollplot <- RockNRoll %>%
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%
ggplot(
  aes(
    x = valence,
    y = danceability,
    color = mode,
    size = energy,
    label = track.name
    )
  ) +
  geom_point(alpha=0.4) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  theme(
    legend.position = "bottom"
  )

ggplotly(rollplot)
```

***
For these visualizations, I used the Spotify playlists as specified before in the introduction. Here we hope to see some of the hypotheses that have been stated earlier.

A very interesting fact that can be concluded from the plots above is that it is apparent that the Rock genre got a lot more diverse as time went on. In the plot of Rock 'n Roll, note that the average danceability and valence are very high. In the period following directly after, the Rock Classics, the valence is a lot more spread out than it was before. The period after this, The Rock Subgenres, it is seen that the average danceability has gone down significantly.

Next to these findings, it is also apparent in the plots that in the Rock 'n Roll period, the overall energy was lower than the energy of later tracks. This can be noted by the fact that the left plot seems less crowded. This effect is however magnified by the fact that the Rock 'n Roll playlist had less songs than the other plots.


### Rock Classics era (70s, 80s)

``` {r}
classicplot <- RockClassics %>%
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%
ggplot(
  aes(
    x = valence,
    y = danceability,
    color = mode,
    size = energy,
    label = track.name
    )
  ) +
  geom_point(alpha=0.4) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  theme(
    legend.position = "bottom"
  )

ggplotly(classicplot)
```

***
Here I will explain what is seen in the plot :)

### Rock Subgenres (90s, 00s)

``` {r}
subgplot <- RockSubGenres %>%
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%
ggplot(
  aes(
    x = valence,
    y = danceability,
    color = mode,
    size = energy,
    label = track.name
    )
  ) +
  geom_point(alpha=0.4) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  theme(
    legend.position = "bottom"
  )

ggplotly(subgplot)
```

***
Here comes the explanation again!!!! (Just not yet)


### Comparing all the different times in one big Rockplot


``` {r}
fullplot <- Rock %>%
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) %>%
ggplot(
  aes(
    x = valence,
    y = danceability,
    color = mode,
    size = energy,
    label = track.name
    )
  ) +
  geom_point(alpha=0.4) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  facet_wrap(~era) +
  theme(
    legend.position = "bottom"
  )

ggplotly(fullplot)

```

*** 
You know it... Here it comes... The anticipation is agonising... Placeholder!!!!!!



